\chapter*{Acknowledgments}\label{sec:acknowledgment}


First of all, I would like to thank my two supervisors Danilo Gligoroski and Colin Boyd for all their guidance and support throughout my studies.
%Danilo was the one that encouraged me to pursue a PhD in cryptography and has always enthusiastically supported me.
%After switching thesis topic to key exchange protocols,
%Colin became my second supervisor and has ever since provided me with all the help and guidance I could need.
%


Much of the work in this thesis is the result of collaboration with others.
I want to thank my co-authors on the two papers on which the main parts of this thesis are based:
Chris Brzuska and Douglas Stebila.
I especially want to acknowledge Chris Brzuska for showing me how fun (and exhausting!) research can be,
but also for being a friend, a mentor, a role-model, and in effect a third supervisor.
Without him this thesis would simply not have been possible.

Additionally,
I want to thank Bogdan Warinschi and Cas Cremers for many helpful discussions.
I am also indebted to those who volunteered their time and effort into proofreading my thesis:
Colin Boyd, Cristina Onete, Chris Brzuska, and Gareth Davies---I express my deepest gratitude to all of you.



A big thanks also goes to my office mates Simona Samardjiska, Britta Hale, and Chris Carr for their great company during my PhD at NTNU.

Finally,
I would like to thank my family for their unwavering support and encouragement throughout the years,
and last but not least, Vilde for always believing in me. Thank you.
